---
title: "Project 1: coursera Reproducable research"

---
###Reading in the data of a fitness tracker
```{r}
dat <-  read.csv("C:\\Users\\Lucas\\Downloads\\repdata_data_activity\\activity.csv",na.strings = c("NA","NaN","","#DIV/0!"))
str(dat)
dat$date <- as.Date(dat$date, format = "%Y-%m-%d")
head(dat)
```

Including packages that are used in the project
```{r}
library(ggplot2)
library(dplyr)
library(mice)
```

### 1. What is the total number of steps taken per day?


Using dplyr, total number of steps is calculated per day using group_by function
Since there are missing values in the dataset, we need to specify na.rm=TRUE
```{r }
sumdat <- dat %>%
            group_by(date) %>%
            summarize(sumStep=sum(steps),na.rm=TRUE)
total_step <- aggregate(steps ~ date, data = dat, sum, na.rm = TRUE)
str(total_step)
```

Chunk of the code used to create a plot 
```{r}
newtheme <- theme_bw() + 
          theme(plot.title = element_text(color = "Black"))

barfill <- "#4271AE"
barlines <- "#1F3552"
```
### 2. Histogram of the total number of steps taken each day?

```{r}
ggplot(sumdat, aes(x=sumStep)) + 
  ggtitle("Total Numer of Steps Taken Each Day") +
  geom_histogram(aes(y = ..count..), binwidth = 500,
                 colour = barlines, fill = barfill)+
  scale_y_continuous(name="Frequency",limits=c(0,8),expand = c(0,0))+
  scale_x_continuous("Steps Count",limits=c(0,22000),expand=c(0,0))+
  theme_bw()+
  theme(panel.grid = element_blank())
```
### 3. Mean and median number of steps taken each day
```{r}
mean(sumdat$sumStep,na.rm=TRUE)
median(sumdat$sumStep,na.rm=TRUE)
```

Preparing data to create average number of steps per interval
```{r}
avgdat <- dat %>%
  group_by(interval) %>%
  summarize(avgStep=mean(steps,na.rm=TRUE))
```

### 4. Time series plot of the average number of steps taken


```{r}
ggplot(avgdat, aes(x=interval,y=avgStep)) + 
  ggtitle("Time Series 5 min. Interval: Average Numer of Steps Taken Each Day") +
  geom_line()+
  scale_y_continuous(name="Average No. of Steps",limits=c(0,200),expand = c(0,0))+
  scale_x_continuous("Interval",limits=c(0,2200),expand=c(0,0))+
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(legend.position = "none") 
```

## 5.The 5-minute interval that, on average, contains the maximum number of steps
```{r}
avgdat$interval[which.max(avgdat$avgStep)]

```
### 6.Code to describe and show a strategy for imputing missing data
```{r}
#Surpressing the output as the code prints all the iterations

sum(is.na(dat))
md.pattern(dat)

imputeDat <- mice(dat, m=2, maxit = 50, method = 'pmm', seed = 500)
```
Referring to documentation:
m  - Refers to 5 imputed data sets
maxit - Refers to no. of iterations taken to impute missing values
method - Refers to method used in imputation. we used predictive mean matching.

```{r}
summary(imputeDat)
head(imputeDat$imp$steps)
completeDat <- complete(imputeDat,2)
```
###7.Histogram of the total number of steps taken each day after missing values are imputed
Preparing the data using imputed values
```{r}
sumCompdat <- completeDat %>%
  group_by(date) %>%
  summarize(sumStep=sum(steps))

str(sumCompdat)

```

###Histogram

```{r}
ggplot(sumCompdat, aes(x=sumStep)) + 
  ggtitle("Total Numer of Steps Taken Each Day (Imputed)") +
  geom_histogram(aes(y = ..count..), binwidth = 500,
                 colour = barlines, fill = barfill)+
  scale_y_continuous(name="Frequency",limits=c(0,8),expand = c(0,0))+
  scale_x_continuous("Steps Count",limits=c(0,22000),expand=c(0,0))+
  theme_bw()+
  theme(panel.grid = element_blank())
```

##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Preparing the data

```{r}

completeDat$day <- weekdays(completeDat$date)
completeDat$week <- ""
completeDat[completeDat$day %in% "Saturday" | completeDat$day %in% "Sunday", ]$week <- "weekend"
completeDat[!(completeDat$day %in% "Saturday" | completeDat$day %in% "Sunday"), ]$week <- "weekday"
completeDat$week <- factor(completeDat$week)

#average using dplyr
avg_step_imp <- completeDat %>%
                  group_by(interval,week) %>%
                  summarize(avgStep=mean(steps))
```

Creating a plot
```{r}


ggplot(avg_step_imp, aes(x=interval,y=avgStep)) + 
  ggtitle("Average Number of Steps Taken (across all weekdays/weekends)") +
  geom_line()+
  facet_grid(week ~ .) +
  scale_y_continuous(name="Average No. of Steps",limits=c(0,200),expand = c(0,0))+
  scale_x_continuous("5 min Interval",limits=c(0,2200),expand=c(0,0))+
  theme_bw()+
  theme(panel.grid = element_blank())+
  theme(legend.position = "none") 
```

```{r}
library(knitr)
knit2html("Project 1.Rmd")
  ```
